---
title: "Project_2_558"
author: "Robbie Ferrand"
format: html
editor: visual
---

Load in all necessary libraries

```{r}
library(tidyverse)
```

Read in the data and view

```{r message = FALSE}
mobile_behavior<- read_csv("user_behavior_dataset.csv")
```

This dataset provides a comprehensive analysis of mobile device usage patterns and user behavior classification. It contains 700 samples of user data, including metrics such as app usage time, screen-on time, battery drain, and data consumption. Each entry is categorized into one of five user behavior classes, ranging from light to extreme usage, allowing for insightful analysis and modeling.

#### Key Features:

-   User ID: Unique identifier for each user.

-   Device Model: Model of the user's smartphone.

-   Operating System: The OS of the device (iOS or Android).

-   App Usage Time: Daily time spent on mobile applications, measured in minutes.

-   Screen On Time: Average hours per day the screen is active.

-   Battery Drain: Daily battery consumption in mAh.

-   Number of Apps Installed: Total apps available on the device.

-   Data Usage: Daily mobile data consumption in megabytes.

-   Age: Age of the user.

-   Gender: Gender of the user (Male or Female).

-   User Behavior Class: Classification of user behavior based on usage patterns (1 to 5). Light to extreme usage.

One-way Contingency Tables

```{r}
table(mobile_behavior$`Operating System`)
table(mobile_behavior$`Device Model`)
table(mobile_behavior$Gender)
table(mobile_behavior$`User Behavior Class`)
```

Two-Way Contingency Tables

```{r}
table(mobile_behavior$`Gender`, mobile_behavior$`Operating System`)

table(mobile_behavior$`User Behavior Class`, mobile_behavior$Gender)

table(mobile_behavior$`User Behavior Class`, mobile_behavior$`Operating System`)
```

off the gate, I think we should be able to subset on Operating System, Gender, and User Behavior Class.

Basic Numeric Summaries

Calculate Min, Max, Mean, Median, SD, IQR for all numeric variables:

```{r}
numerical_summaries <- mobile_behavior |>
  summarise(across(where(is.numeric), 
  list(
        Min = ~min(. , na.rm = TRUE),
        Max = ~max(. , na.rm = TRUE),
        Mean = ~mean(. , na.rm = TRUE),
        Median = ~median(. , na.rm = TRUE),
        SD = ~sd(. , na.rm = TRUE),
        IQR = ~IQR(. , na.rm = TRUE)), 
        .names = "{.col}_{.fn}")) |>
  pivot_longer(everything(), names_to = c("Variable", ".value"), 
               names_sep = "_") |>
  #slice allows us to remove rows! it's like select but for rows
  slice(-c(1, 8))
```
